---
title: "CEM Notes"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
install.packages("cem")
library(cem)

library(tidyverse)
```

Use the National Supported Work Demonstration data (Lalonde data set)
treatment variable -- **treated** is 1 for the treatment group and 0 for the control group
outcome variable -- **re78**; earnings in 1978
Pre-treatment variables to control for:
* age; **age**; categorical
* years of education; **education**; categorical
* marital status; **married**; binary
* lack of a high school diploma; **nodegree**; binary
* race; **black, hispanic**; binary
* unemployment in 1974 and 1975; **u74**, **u75**; binary
* real earnings in 1974 and 1975; **re74**, **re75**; continuous and highly skewed
* **q1**; a survey question asking an opinion about this job training program; strongly agree, agree, neutral, strongly disagree, disagree, and no options
  -- Added by Iacus, King, & Porro (2020)

```{r}
require(cem)
data("LeLonde")

# Remove missing data
Le <- data.frame(na.omit(LeLonde))

# Compare the side of the treated and control groups
tr <- which(Le$treated == 1)
ct <- which(Le$treated == 0)
ntr <- length(tr)  # 258 treated units
nct <- length(ct)  # 392 control units

# Unadjusted difference in means is 759
mean(Le$re78[tr]) - mean(Le$re78[ct])

# pre-treatment variables to control for
vars <- c("age", "education", "black", "married", "nodegree", "re74", "re75", "hispanic", "u74", "u75","q1")

# Compute imbalance 
imbalance(group = Le$treated, data = Le[vars])

# Drop some variables before computation
todrop <- c("treated", "re78")
imbalance(group = Le$treated, data = Le, drop = todrop)

## Automated coarsening
# match on all variables but re78 because that's the outcome variable
mat <- cem(treatment = "treated", data = Le, drop = "re78", keep.all = TRUE)
mat

## Choosing the coarsening yourself, rather than using the automated algorithm
# For categorical variables, we use the grouping option
# group q1 into three groups
q1.grp <- list(c("strongly agree", "agree"), c("neutral", "no opinion"), c("strongly disagree", "disagree"))

# For numerical variables, we use the cutpoints option
# group years of education into grade school, middle school, high school, college, and graduate school
# 0-6, 708, 9-12, 13-16, > 16
educut <- c(0, 6.5, 8.5, 12.5, 17)

# run cem again
mat1 <- cem(treatment = "treated", data = Le, drop = "re78",
            cutpoints = list(education = educut), grouping = list(q1 = q1.grp))
mat1

## Progressive coarsening
# try relax the cem solution by hand
# a is to have 10 bins on age
a <- cem("treated", Le, cutpoints = list(age = 10), drop = "re78", grouping = list(q1 = q1.grp))
# b -- 6 bins on age
b <- cem("treated", Le, cutpoints = list(age = 6), drop = "re78", grouping = list(q1 = q1.grp))
# c -- 3 bins on age
c <- cem("treated", Le, cutpoints = list(age = 3), drop = "re78", grouping = list(q1 = q1.grp))
# fewer cutpoints (larger bins) produces more matched units and higher maximum

tab <- relax.cem(mat, Le, depth = 1, perc = 0.3) # only the solutions with at least 30% of the units are matched

plot(tab, group = "1", perc = 0.35, unique = TRUE)

# k-to-k match; the same number of treated and control units in all strata
mat <- cem(treatment = "treated", data = Le, drop = "re78", keep.all = TRUE)
mat$k2k

mat2 <- k2k(mat, Le, "euclidean", 1)
mat2$k2k

# Estimate the causal effect from cem output
data(LL)
head(LL)
mat <- cem(treatment = "treated", data = LL, drop = "re78")
est <- att(mat, re78 ~ treated, data = LL) # linear model
est

# To apply a statistical model to control for the remaining imbalance
est2 <- att(mat, re78 ~ treated + re74, data = LL)
est2

# Dealing with missing values
# Matching on missingness (treat missing values as one of the values of the variables)
mat3 <- cem("treated", LeLonde, drop = "re78", cutpoints = mat$breaks, grouping = list(q1 = q1.grp))

# Matching multiply imputed data
summary(LeLonde)


install.packages("Amelia") # to create multiply imputed datasets
library(Amelia)
require(Amelia)

imputed <- amelia(LeLonde, noms = c("black", "hispanic", "treated", "married", "nodegree", "u74", "u75", "q1"))

mat2 <- cem("treated", datalist = imputed, drop = "re78", data = LeLonde, grouping = list(q1 = q1.grp))

# Create paired sample
data(LL)
head(LL)
mat <- cem("treated",data = LL, drop = "re78")
psample <- pair(mat, data = LL)
```

